Accelerating Neural Network Training using Arbitrary Precision Approximating Matrix Multiplication Algorithms

### 2021-04-24 latex/ICPP21/paper.tex
update paper on vgg figures

Our contribution is the use of APA matrix multiplication algorithms to address this problem.
In particular we
\begin{enumerate}
	\item curate a collection of both well-established and recently discovered practical APA algorithms;
	\item extend the framework of \cite{BB15} to generate efficient multithreaded code for all of them, achieving up to 21\% performance improvement over the best parallel classical implementation;
	\item demonstrate the robustness of learning accuracy to approximate matrix multiplications;
	\item present multithreaded performance improvements of a synthetic MLP of up to 13\% over the use of classical matrix multiplication; and
	\item \RZ{present performance improvements of VGG-11? network of ?\% over the classical matrix multiplication.}
\end{enumerate}

@@

\begin{table}
\centering
\caption{Properties of APA algorithms. Speedup and error are computed assuming 1 recursive step}
\label{tab:algs}
\begin{tabular}{| c | c c c | c c c |} 
\hline
\textbf{Ref} & \textbf{Dims} & \textbf{Rank} & \textbf{Speedup} & $\mathbf{\sigma}$ & $\mathbf{\varphi}$ & \textbf{Error} \\
\hline
- & $\dims{2,2,2}$ & 8 & - & 1 & 0 & \num{1.2e-7} \\
\hline
\cite{BCRL79} & $\dims{3,2,2}$ & 10 & $20\%$ & 1 & 1 & \num{3.5e-4} \\
\cite{AS13} & $\dims{4,2,2}$ & 13 & $23\%$ & 1 & 2 & \num{4.9e-3} \\
\cite{Smirnov13} & $\dims{3,3,2}$ & 14 & $29\%$ & 1 & 3 & \num{1.9e-2} \\
\cite{Smirnov13} & $\dims{5,2,2}$ & 16 & $25\%$ & 1 & 3 & \num{1.9e-2} \\
\cite{Smirnov13} & $\dims{3,3,3}$ & 20 & $35\%$ & 1 & 6 & \num{1.0e-1} \\
\cite{Schonhage81} & $\dims{3,3,3}$ & 21 & $29\%$ & 1 & 2 & \num{4.9e-3} \\
\cite{Smirnov15} & $\dims{7,2,2}$ & 22 & $27\%$ & 1 & 5 & \num{7.0e-2} \\
\cite{Smirnov16b} & $\dims{4,4,2}$ & 24 & $33\%$ & 1 & 3 & \num{1.9e-2} \\
\cite{Smirnov16a} & $\dims{4,3,3}$ & 27 & $33\%$ & 1 & 3 & \num{1.9e-2} \\
\cite{Smirnov16b} & $\dims{5,5,2}$ & 37 & $35\%$ & 1 & 3 & \num{1.9e-2} \\
\cite{Smirnov14} & $\dims{4,4,4}$ & 46 & $39\%$ & 1 & 3 & \num{1.9e-2} \\
\cite{Smirnov18} & $\dims{5,5,5}$ & 90 & $39\%$ & 1 & 3 & \num{1.9e-2} \\
\hline
\end{tabular}
\end{table}

@@

\section{Application of APA algorithms in VGG-11}

\begin{figure}[hb!]
\centering
\input{plots/VGG_time}
\caption{Total time test on VGG-11 with 6 threads}
\label{fig:vgg}
\end{figure}

\subsection{VGG-11 Network}
\RZ{description of the model of VGG-11 and VGG-19?}

@@

\subsection{Performance}
\RZ{VGG-11 fully connected layer: BS x 25K x 4K / BS x 4K x 4K / BS x 4K x 1K? with plot}
\cref{fig:vgg}
\RZ{Fast Combo: 525, 442, 242}

@@

While the accuracy and performance results are promising for MLP networks and the MNIST dataset, we believe that APA algorithms can also be effective in state-of-the-art networks and more difficult machine learning problems.
For example, the VGG-19 network for large-scale visual recognition \cite{SZ15} is based on a deep convolutional network and is bottlenecked by large fully connected layers, including a layer with dimensions $25088\times 4096$, where APA algorithms can offer significant performance improvement.

@@

\newpage
\balance
\bibliographystyle{ACM-Reference-Format}
\bibliography{refs}

>>>>

Our contribution is the use of APA matrix multiplication algorithms to address this problem.
In particular we
\begin{enumerate}
	\item curate a collection of both well-established and recently discovered practical APA algorithms;
	\item extend the framework of \cite{BB15} to generate efficient multithreaded code for all of them, achieving up to 21\% performance improvement over the best parallel classical implementation;
	\item demonstrate the robustness of learning accuracy to approximate matrix multiplications;
	\item present multithreaded performance improvements of a synthetic MLP of up to 13\% over the use of classical matrix multiplication; and
	\item \RZ{present performance improvements of VGG-19 network of 15\% over the classical matrix multiplication.}
\end{enumerate}

@@

\begin{table}
\centering
\begin{tabular}{| c | c c c | c c c |} 
\hline
\textbf{Ref} & \textbf{Dims} & \textbf{Rank} & \textbf{Speedup} & $\mathbf{\sigma}$ & $\mathbf{\varphi}$ & \textbf{Error} \\
\hline
- & $\dims{2,2,2}$ & 8 & - & 1 & 0 & \num{1.2e-7} \\
\hline
\cite{BCRL79} & $\dims{3,2,2}$ & 10 & $20\%$ & 1 & 1 & \num{3.5e-4} \\
\cite{AS13} & $\dims{4,2,2}$ & 13 & $23\%$ & 1 & 2 & \num{4.9e-3} \\
\cite{Smirnov13} & $\dims{3,3,2}$ & 14 & $29\%$ & 1 & 3 & \num{1.9e-2} \\
\cite{Smirnov13} & $\dims{5,2,2}$ & 16 & $25\%$ & 1 & 3 & \num{1.9e-2} \\
\cite{Smirnov13} & $\dims{3,3,3}$ & 20 & $35\%$ & 1 & 6 & \num{1.0e-1} \\
\cite{Schonhage81} & $\dims{3,3,3}$ & 21 & $29\%$ & 1 & 2 & \num{4.9e-3} \\
\cite{Smirnov15} & $\dims{7,2,2}$ & 22 & $27\%$ & 1 & 5 & \num{7.0e-2} \\
\cite{Smirnov16b} & $\dims{4,4,2}$ & 24 & $33\%$ & 1 & 3 & \num{1.9e-2} \\
\cite{Smirnov16a} & $\dims{4,3,3}$ & 27 & $33\%$ & 1 & 3 & \num{1.9e-2} \\
\cite{Smirnov16b} & $\dims{5,5,2}$ & 37 & $35\%$ & 1 & 3 & \num{1.9e-2} \\
\cite{Smirnov14} & $\dims{4,4,4}$ & 46 & $39\%$ & 1 & 3 & \num{1.9e-2} \\
\cite{Smirnov18} & $\dims{5,5,5}$ & 90 & $39\%$ & 1 & 3 & \num{1.9e-2} \\
\hline
\end{tabular}
\caption{Properties of APA algorithms. Speedup and error are computed assuming 1 recursive step}
\label{tab:algs}
\end{table}

@@

\section{Application of APA algorithms in VGG-19}

\begin{figure*}
%
\subfloat[One thread]{
\centering
\input{plots/VGG_time_seq}
\label{fig:vgg_seq}}
	%
\hfill	
\subfloat[Six threads]{
\centering
\input{plots/VGG_time}
\label{fig:vgg_par6}}
%
\caption{Total time test on the fully connected layes on VGG-19 using $\dims{4,4,2}$ and dgemm}
\label{fig:vgg}
\end{figure*}

\subsection{VGG-19 Network}
VGG19 is a convolutional neural network typically trained on the imagenet dataset as a classification task. The model is one of the highest performing algorithms for this task. 
The network is composed of 19 layers, 16 convolutional and 3 fully connected. 
The 3 fully connected layers consist of 25,088, 4096 and 1000 nodes and are placed at the end of the network. 
These create very large multiplications in forwards and backwards propagation. 
Using faster APA algorithms for the multiplication we are able to speed up the training of the network.

@@

\subsection{Performance}

Replacing the classic matrix multiplication algorithm in these layers with the Smirnov 442 algorithm provides up to a 10\% speed up of the fully connected layers during training with 6 threads and a 15\% speed up using a single thread. 
\cref{fig:vgg} plot the training time of each network with increasing batch sizes

@@

While the accuracy and performance results are promising for MLP networks and the MNIST dataset, we believe that APA algorithms can also be effective in state-of-the-art networks and more difficult machine learning problems.
For example, the VGG-19 network for large-scale visual recognition \cite{SZ15} is based on a deep convolutional network and is bottlenecked by large fully connected layers.
Our result are able to show a 15\% performance improvement using the sequential algorithm, and a 10\% improvement using parallel algorithm.
This highlights the practical applications on training large networks.

@@

\balance
\bibliographystyle{ACM-Reference-Format}
\bibliography{refs}
